---
title: "FLT3-ITD Cooperating Mutations Unsupervised Clustering"
author: "Jenny Smith"
date: "9/28/21"
output: html_document
---

# Set-up 

```{r setup, cache = FALSE, include = FALSE}
require(knitr)
knitr::opts_knit$set(root.dir = file.path(PROJHOME,"2021.04.01_FLT3.ITD_Cooperating_Mutations/"))
```

```{r}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),
                      tidy=TRUE,
                      fig.align='center',
                      fig.width = 10, fig.height = 10)
node=Sys.info()[["nodename"]]
if(!grepl("local", node)){
  print(node)
  options(bitmapType = 'cairo')
  grDevices::X11.options(type='cairo')
}

options(stringsAsFactors = FALSE)
table = function (..., useNA = 'ifany') base::table(..., useNA = useNA)
```

```{r message = FALSE, warning=FALSE}
library(stringr)
library(magrittr)

library(ggplot2)
library(gridExtra)
library(patchwork)
library(ggpubr)

library(dplyr)
library(tidyr)
library(tibble)
library(REDCapR)

library(DeGSEA)
library(edgeR)
library(DESeq2)

getwd()
```

# Define Functions

```{r}
make_alpha <- function(color, percent=100){
  # https://www.dataanalytics.org.uk/make-transparent-colors-in-r/
  ## Get RGB values for named color
  rgb.val <- col2rgb(color)
  
  ## Make new color using input color as base and alpha set by transparency
  t.col <- rgb(rgb.val[1], rgb.val[2], rgb.val[3],
               max = 255,
               alpha = (100 - percent) * 255 / 100)

}
```

```{r}
#https://stackoverflow.com/questions/9253843/r-clustering-purity-metric
#https://stats.stackexchange.com/questions/95731/how-to-calculate-purity
calc_cluster_purity <- function(clusters, classes) {
  assertthat::assert_that(length(clusters)==length(classes),
                          msg="Length of clusters is not equal to lenght of classes vector.")
  
  confusion_mat <- table(classes, clusters)
  max_per_cluster <- apply(confusion_mat, 2, max)
  total <- length(clusters)
  sum(max_per_cluster) / total
}
```

```{r}
UMAP_function <- function(data_matrix,
                          metadata,N,
                          y=NULL,
                          pca=NULL,
                          target_metric="euclidean",
                          min_dist=0.1,n_neighbors=15L,
                          scale_data=FALSE, threads=2){
  #data_matrix is a expression matrix, tfidf transformed with samples as columns, genes as rows. It has been subset to selected genes.
  set.seed(2020)
  UMAP<-uwot::umap(X = t(data_matrix), 
                   n_components = N,
                   scale = scale_data,
                   pca=pca,
                   y=y,
                   target_metric = target_metric,
                   metric = "cosine",
                   spread= 1,
                   min_dist = min_dist, #originally 0.1 by Scott Furlan
                   n_neighbors = n_neighbors, #was default 15 
                   fast_sgd = FALSE, 
                   n_epochs=500, #increased from default 200
                   
                   init = "spectral", #default
                   nn_method = "annoy", #Use approximate nearest neighbors 
                   n_trees=200, #default is 50
                   search_k=5*n_neighbors*200, #default is 2*n_neighbors*n_trees
                   repulsion_strength=1, #default
                   
                   n_threads=threads,
                   tmpdir=file.path(SCRATCH,"jlsmith3"),
                   ret_model=TRUE, #for new data with umap_transform()
                   ret_nn=TRUE,
                   verbose=T)
  gc()
  toplot<-data.frame(x=UMAP$embedding[,1],y=UMAP$embedding[,2])
  if(ncol(UMAP$embedding)==3){toplot[["z"]] <- UMAP$embedding[,3]}
  toplot<-as.data.frame(cbind(toplot, metadata))
  gc()
  
  res <- list("umap_df"=toplot, "model"=UMAP)
  
  return(res)
}
```

# Read in the counts

```{r}
genome <- "GRCh38"
```

## GRCh38 

```{r}
current_files <- dir(file.path(PROJHOME, "0000.00.03_ExpressionMatrices/Kallisto_GRCh38_Gencode_v29/"))
# current_files
    

if(genome=="GRCh38"){
    cts_grch38_file <- grep("_RBD_.+scaledTPM_counts.RDS", current_files, value=TRUE)
    cts_grch38 <- readRDS(file.path(PROJHOME, "0000.00.03_ExpressionMatrices/Kallisto_GRCh38_Gencode_v29/",cts_grch38_file))
    cts_grch38_ids <- cts_grch38[,grep("gene_id|gene_name", colnames(cts_grch38))]
    
    cts_grch38 <- as.data.frame(cts_grch38)
    rownames(cts_grch38) <-  cts_grch38_ids$gene_name
    cts_grch38 <- cts_grch38[,-grep("gene_id|gene_name", colnames(cts_grch38))]
    
    # head(cts_grch38[,1:5])
    dim(cts_grch38) #58263  3021 
    
    ### TPM
    TPM_grch38_file <- grep("_RBD_.+Abundance_TPM", current_files, value=TRUE)
    TPM_grch38 <- readRDS(file.path(PROJHOME, "0000.00.03_ExpressionMatrices/Kallisto_GRCh38_Gencode_v29/",TPM_grch38_file))
    TPM_grch38_ids <- TPM_grch38[,grep("gene_id|gene_name", colnames(TPM_grch38))]
    
    TPM_grch38 <- as.data.frame(TPM_grch38)
    rownames(TPM_grch38) <-  TPM_grch38_ids$gene_id
    TPM_grch38 <- TPM_grch38[,-grep("gene_id|gene_name", colnames(TPM_grch38))]
    
    
    # head(TPM_grch38)
    dim(TPM_grch38) #58263  3021
}
```


# ClinData

```{r message=FALSE}
#https://cran.r-project.org/web/packages/REDCapR/vignettes/workflow-read.html 
project <- "FLT3-ITD_UBTF-ITD_KMT2A-PTD_Cooperating_Mutations"

if(project==""){
  stop("Must include Project name!")
  
}else{
  current_cde_database <- paste("TARGET_AML_CDEs_For_Project",project, ".RDS", sep="_")
  
  if(file.exists(current_cde_database)){
    merged <- readRDS(current_cde_database)
      
  }else{
    path_credential <- file.path(HOME,".redcap")
    project_id <- 1295
  
    credential  <- REDCapR::retrieve_credential_local(
      path_credential = path_credential,
      project_id = project_id)
    
    #takes about 30 sec to download. 
    merged <- REDCapR::redcap_read(redcap_uri = credential$redcap_uri, 
                          token = credential$token, 
                          raw_or_label_headers = 'label')
    if(merged$success){
      merged <- data.frame(merged$data, check.names=TRUE) #remove the white spaces  (will this even work??)
      saveRDS(merged, current_cde_database)
    }
    
    #Create a simple log file from the day the project starts
    cat(c(paste("Date:", Sys.Date()),
          paste("cts:", cts_grch38_file),
          paste("tpm:", TPM_grch38_file), 
          paste("CDE:", current_cde_database)),
          sep = "\n", 
          file = paste(project, Sys.Date(), ".log", sep="_"))
    
  }
  
  #keep a list of the ineligable patiens to remove if necessary
  inelig <- merged %>% 
    filter(Eligibility.Comments == "remove") %>% 
    pull(USI)
  
  #Filter those who are eligible for the study
  merged <- merged %>% 
    filter(Eligibility.Comments != "remove")
  
  
  dim(merged) #3526  156
  # head(merged)
}
```

```{r}
sample_info <- read.csv(file.path(TARGET, "SequencingDataMatrix/TARGET_AML_Ribodepleted_Manifest_08.12.21.csv"))  %>% 
  filter(!USI %in% inelig)

dim(sample_info)
```

```{r}
orig <- read.csv(file.path(CDE,"Merged/TARGET_AML_0531_1031_merged_CDEs_05.21.21.csv")) %>% 
         filter(Eligibility_Comments != "remove")

dim(orig)


#### OKAY ####
# check <- orig %>% 
#   select(USI, FLT3.ITD.positive., NPM.mutation., CEBPA.mutation., WT1.mutation.) %>% 
#   left_join(., select(merged, USI,  FLT3.ITD.positive., NPM.mutation., CEBPA.mutation., WT1.mutation.),
#             by="USI") %>% 
#   select(colnames(.)[order(colnames(.))])


# check
# identical(check$FLT3.ITD.positive..x, check$FLT3.ITD.positive..y)
# identical(check$NPM.mutation..x, check$NPM.mutation..y)
# identical(check$CEBPA.mutation..x, check$CEBPA.mutation..y)
# identical(check$WT1.mutation..x, check$WT1.mutation..y)
```

```{r}
cohort <- read.csv("pts_3033_for_FLT3_Manuscript_Regs_COG.csv") %>% 
  # openxlsx::read.xlsx("pts_3033_for_FLT3_Manuscript_Regs_COG.xlsx") %>%
  mutate_all(as.character) %>% 
  left_join(., select(orig,Reg., USI), by=c("reg_no"="Reg.")) %>% 
  arrange(USI)

head(cohort)
# dim(cohort) #3033    1
# any(is.na(cohort$USI))
# View(cohort)
```



# Select Samples

(If you’re familiar with quote() and substitute() in base R, quo() is equivalent to quote() and enquo() is equivalent to substitute().)

```{r}
#A function to make the mutation columns contain the mutation name instead of  "Yes"
convert_yes <- function(mut_column){
  
  # mut <- rlang::ensym(mut_column) #evaluated as .x
  # mut <- enquo(mut_column) #evaluated as .x
  # mut <- quo(mut_column) #evaluated as mut_column
  # mut <- eval.parent(enquo(mut_column), n = 2) #evaluated as .x 
  # print(mut)

  mut <- substitute(mut_column)
  mode(mut) <- "character"

  gsub("Yes", mut, mut_column) %>%
    gsub(".positive.|.mutation.", "+", .) %>%
    gsub("No$|None$|Unknown", NA_character_, .)
  
}
```

```{r}
samples <- sample_info %>% 
  left_join(., select(merged, USI, FLT3.ITD.positive.,
                      UBTF.ITD.positive.,KMT2A.PTD.positive.,
                      WT1.mutation.,
                      NPM.mutation.,CEBPA.mutation., trisomy.8, ISCN,
                      Mutations.Category,EFS.event.type.ID,
                      matches("^EFS|^Event|^OS")), 
            by="USI")  %>% 
  filter(grepl("^AML$|NBM|CD34_PB", Group), Time_point != "AML") %>% 
  filter(Group=="NBM" |  Group == "CD34_PB" | USI %in% cohort$USI) %>% 
  filter(!grepl("_replicate", Sample)) %>% 
  filter(Sample %in% colnames(cts_grch38)) %>% 

  mutate_at(vars(Time_point), ~factor(.,
                                      levels=c("diagnostic","relapse","remission",
                                               "NBM","CD34_PB"))) %>%
  mutate_at(vars(Additional.Fusions.CNV), ~ifelse(.=="Unknown", NA, .)) %>%
  mutate_at(vars(trisomy.8), ~case_when(
    .=="Unknown" & ISCN != "Unknown" ~ NA_character_,
    TRUE ~ .)) %>%
  arrange(Time_point, desc(FLT3.ITD.positive.)) %>%

  #Create a column with the co-occuring mutations
  #worked once a while ago and doesnt work anymore using across()
  # mutate(across(.cols = c(FLT3.ITD.positive.,WT1.mutation., NPM.mutation.,CEBPA.mutation.,
  #                         UBTF.ITD.positive.,KMT2A.PTD.positive.),
  #               .fns = ~convert_yes(.x),
  #               .names="{.col}_edit")) %>%
  # 
  mutate(FLT3.ITD.positive._edit=convert_yes(FLT3.ITD.positive.),
         WT1.mutation._edit=convert_yes(WT1.mutation.),
         NPM.mutation._edit=convert_yes(NPM.mutation.),
         CEBPA.mutation._edit=convert_yes(CEBPA.mutation.),
         UBTF.ITD.positive._edit=convert_yes(UBTF.ITD.positive.),
         KMT2A.PTD.positive._edit=convert_yes(KMT2A.PTD.positive.)) %>%

  #clean up the Primary.Fusion to convert "None" to NA
  mutate(Primary.Fusion2=gsub("None|Unknown",NA_character_, Primary.Fusion)) %>%
  unite(col="FLT3.Groups",
        Primary.Fusion2, matches("^(FLT3|WT1|NPM|CEBPA).+edit"),
        sep="/", remove=FALSE, na.rm=TRUE) %>%
  unite(col="FLT3.UBTF.KMT2A_Groups",
        matches("^(FLT3|UBTF|KMT2A).+edit"),
        sep="/", remove=TRUE, na.rm=TRUE) %>%

  #Clean up the groups to retain only those of high risk (poor risk)  for the plots
  mutate(FLT3.Groups_PoorRisk=case_when(
    #Define poor risk groups
    grepl("^NUP98-NSD1\\/FLT3", FLT3.Groups,fixed = F) ~ FLT3.Groups,
    grepl("^FLT3.ITD\\+\\/WT1\\+$", FLT3.Groups, fixed = F) ~ FLT3.Groups,
    grepl("FLT3.ITD\\+\\/WT1\\+", FLT3.Groups) ~ "FLT3.ITD+/WT1+",
    grepl("DEK-NUP214\\/FLT3.ITD\\+$", FLT3.Groups, fixed = F) ~ FLT3.Groups,

    #Define favorable co-occuring muts
    grepl("FLT3.ITD\\+\\/NPM\\+$|FLT3.ITD\\+\\/CEBPA\\+$|CBFB-MYH11\\/FLT3.ITD\\+$|RUNX1-RUNX1T1\\/FLT3.ITD\\+$",
          FLT3.Groups, fixed = F) ~ "Low risk/FLT3-ITD+",

    grepl("FLT3.ITD\\+", FLT3.Groups) ~ "All others/FLT3.ITD+",
    TRUE ~ AML_Subtype)) %>%


  #Clean up the groups to retain those with favorable (low risk) risk for the plots
  mutate(FLT3.Groups_FavRisk=case_when(
    #Poor risk
    grepl("NUP98-NSD1\\/FLT3", FLT3.Groups,fixed = F) ~ "High risk/FLT3-ITD+",
    grepl("^FLT3.ITD\\+\\/WT1\\+$", FLT3.Groups, fixed = F) ~ "High risk/FLT3-ITD+",
    grepl("FLT3.ITD\\+\\/WT1\\+", FLT3.Groups) ~ "High risk/FLT3-ITD+",
    grepl("DEK-NUP214\\/FLT3.ITD\\+$", FLT3.Groups, fixed = F) ~ FLT3.Groups,

     #Define favorable co-occuring muts
    grepl("FLT3.ITD\\+\\/NPM\\+$|FLT3.ITD\\+\\/CEBPA\\+$", FLT3.Groups, fixed = F) ~ "NPM+ or CEBPA+/FLT3-ITD+",
    grepl("CBFB-MYH11\\/FLT3.ITD\\+$", FLT3.Groups) ~ FLT3.Groups,
    grepl("RUNX1-RUNX1T1\\/FLT3.ITD\\+$", FLT3.Groups) ~ FLT3.Groups,

    grepl("FLT3.ITD\\+", FLT3.Groups) ~  "All others/FLT3.ITD+",
    TRUE ~ AML_Subtype)) %>%

  mutate_at(vars(FLT3.UBTF.KMT2A_Groups), ~case_when(
    .=="" & grepl("CBFB-MYH11|KMT2A|RUNX1-RUNX1T1", AML_Subtype) ~paste0(AML_Subtype,"/FLT3-ITD-"),
    .=="" & grepl("NBM|CD34_PB", AML_Subtype) ~ AML_Subtype,
    .=="" ~ "All others/FLT3.ITD-",
    TRUE ~ .)) %>%

  mutate_at(vars(Batch), ~gsub("rlps[1-4]", "rlps", .)) %>%
  mutate(USI1=USI,
         USI=Sample) %>%

  select(USI, Sample, Protocol,Time_point,
         matches("FLT3.Groups"),
         FLT3.ITD.positive.,Primary.Fusion, WT1.mutation.,
         NPM.mutation.,CEBPA.mutation.,trisomy.8, Primary.Fusion,
         Additional.Fusions.CNV,ISCN,
         everything())
  


# head(samples)
dim(samples) # 2284   28
# write.csv(select(samples,-Reg.), "FLT3.ITD_Manuscript_Cohort_with_RNAseq_mutations_and_fusion_info.csv", row.names = FALSE)
```

```{r}
diagnostic_samples <- samples %>% 
  filter(grepl("NBM|CD34_PB|diagnostic", Time_point)) %>% 
  
  group_by(FLT3.Groups_PoorRisk, Batch) %>%
  mutate(Batch_noSingletons=case_when(
    n() < 2 ~ "dx2",
    TRUE ~ Batch)) %>%
  ungroup() %>%
  mutate_at(vars(FLT3.Groups_PoorRisk, FLT3.Groups_FavRisk), ~case_when(
    FLT3.ITD.positive. == "Yes" ~ .,
    grepl("NBM|CD34", Group) ~ Group,
    TRUE ~ "FLT3-ITD-")) %>%
  as.data.frame() %>%
  
  mutate(FLT3.ITD_Groups_Simple=case_when(
    grepl("Low risk", FLT3.Groups_PoorRisk) ~ FLT3.Groups_PoorRisk, 
    grepl("High risk", FLT3.Groups_FavRisk) ~ FLT3.Groups_FavRisk,
    grepl("DEK-NUP214",FLT3.Groups_FavRisk) ~ "High risk/FLT3-ITD+",
    TRUE ~ FLT3.Groups_FavRisk)) %>% 

  mutate_at(vars(AML_Subtype), ~case_when(
    .=="ETS-Fusion" | .=="RBM15-MKL1" ~ "AML, NOS",
    TRUE ~ .)) %>%
  droplevels() %>% 
  set_rownames(.$Sample)


dim(diagnostic_samples)
# table(diagnostic_samples$Group, diagnostic_samples$FLT3.ITD.positive.)
# table(diagnostic_samples$AML_Subtype)
# table(diagnostic_samples$FLT3.Groups_PoorRisk)
# table(diagnostic_samples$FLT3.Groups_FavRisk)
# table(diagnostic_samples$FLT3.UBTF.KMT2A_Groups)

# table(diagnostic_samples$Protocol)
```

```{r}
FLT3.ITD_only <- diagnostic_samples %>% 
  filter(FLT3.ITD.positive.=="Yes") %>% 
  droplevels() %>% 
  set_rownames(.$Sample)  



dim(FLT3.ITD_only)
# head(FLT3.ITD_only)
# table(FLT3.ITD_only$FLT3.Groups, FLT3.ITD_only$Batch)
# table(FLT3.ITD_only$FLT3.Groups_FavRisk)
# table(FLT3.ITD_only$FLT3.Groups_PoorRisk)
# table(FLT3.ITD_only$FLT3.ITD_Groups_Simple)
```


```{r}
# write.csv(select(diagnostic_samples, Sample,FLT3.ITD.positive., matches("FLT3.Groups")) %>%
#             filter(FLT3.ITD.positive.=="Yes") %>%
#             arrange(FLT3.Groups),
#           "TARGET_AML_FLT3-ITD_Cooperating_Mutations_RNAseq_Samples_10.13.2021.csv", row.names = FALSE)
```


# Subset Counts

```{r}
in_counts <- cts_grch38[,diagnostic_samples$Sample]

dim(in_counts) # 58263  1551
head(in_counts[,1:5])
```

```{r}
AML <- ! grepl("BM[0-9]|R[O0][0-9]", colnames(in_counts))
keep <- rowSums(cpm(in_counts[,AML]) >= 1) >= 0.01*ncol(in_counts[,AML])
cts.filtered <- in_counts[keep, ]

dge <- DGEList(counts=cts.filtered)
dge <- calcNormFactors(dge,method = "TMMwsp")

logCPM <- edgeR::cpm(dge,log=TRUE,normalized.lib.sizes=TRUE, prior.count=1)
CPM <- edgeR::cpm(dge,log=FALSE,normalized.lib.sizes=TRUE, prior.count=1)

dim(logCPM) #30690  1551
head(logCPM[,1:5])
```

```{r}
cts.filtered.ids <- cts_grch38_ids %>% 
  filter(gene_name %in% rownames(cts.filtered)) %>% 
  filter(grepl("ENSG", gene_id))

# Im wondering if I want to do VST on only FLT3-ITD or the whole dataset to capture variablility of FLT3-ITD in the presence of other AML subtypes?
# or do I want to really look at intra-FLT3-ITD variablility?
dds <- DESeq2::DESeqDataSetFromMatrix(round(cts.filtered[cts.filtered.ids$gene_name, ], digits = 0),
                                      colData = diagnostic_samples,
                                      design = ~ 1)

# normFactors <- normFactors / exp(rowMeans(log(normFactors)))
# normalizationFactors(dds) <- normFactors
dds <- DESeq2::estimateSizeFactors(dds)
dds_cts <- DESeq2::counts(dds, normalized=FALSE)

# plotDispEsts(dds)
vst <- DESeq2::vst(dds, blind = TRUE)
dim(vst)
```



# Unsupervised Clustering

## 1031 UBTF-ITD and KMT2A-PTD

### All Diagnostic 1031: UBTF-ITD and KMT2A-PTD Focus

```{r}
diagnosic_1031 <- diagnostic_samples %>% 
  filter(grepl("AAML1031",Protocol) | grepl("NBM|CD34_PB", Group))


# table(diagnosic_1031$Protocol)
# table(diagnosic_1031$FLT3.UBTF.KMT2A_Groups)
# diagnosic_1031 %>% 
#   filter(USI1 %in% check_usi$USI) %>% 
#   select(Sample,FLT3.UBTF.KMT2A_Groups)
```

```{r}
library(viridis)
library(RColorBrewer)

Cols <- c("FLT3.Groups_PoorRisk",
          "AML_Subtype",
          "Batch","Tissue")

# all(Cols %in% colnames(FLT3.ITD_only)) #TRUE


pal <- get_palette("npg", 6)

cc_heatmap <- list()
# cc_heatmap[["Batch"]] <- viridis(n=length(unique(FLT3.ITD_only$Batch))) %>% 
#   set_names(unique(FLT3.ITD_only$Batch))
# 
# cc_heatmap[["Tissue"]] <- viridis(n=length(unique(FLT3.ITD_only$Tissue))) %>% 
#   set_names(unique(FLT3.ITD_only$Tissue))

# cc_heatmap[["AML_Subtype"]] <- brewer.pal(length(unique(FLT3.ITD_only$AML_Subtype)),"Paired") %>% 
#   set_names(unique(FLT3.ITD_only$AML_Subtype))

simple_pal <- get_palette("jco", 3)
cc_heatmap[["FLT3.ITD_Groups_Simple"]] <-  simple_pal %>% 
  set_names(c("High risk/FLT3-ITD+", "Low risk/FLT3-ITD+","All others/FLT3.ITD+"))

cc_heatmap[["FLT3.Groups_PoorRisk"]] <- c(
  "All others/FLT3.ITD+"=pal[6],
  "DEK-NUP214/FLT3.ITD+"=pal[3],
  "FLT3.ITD+/WT1+"=pal[1],
  "Low risk/FLT3-ITD+"="#EFC000FF",
  "NUP98-NSD1/FLT3.ITD+"=pal[2],
  "NUP98-NSD1/FLT3.ITD+/WT1+"=pal[4]
)

par(mar=c(15,5,5,5))
input <- cc_heatmap$FLT3.ITD_Groups_Simple
barplot(rep(1,length(input)),
        col = input,
        names.arg = names(input),
        las=2)

# lapply(names(cc_heatmap), function(x) table(FLT3.ITD_only[[x]]))
# lapply(names(cc_heatmap), function(x) names(cc_heatmap[[x]])[!names(cc_heatmap[[x]]) %in% FLT3.ITD_only[[x]]])
```

```{r}
library(DelayedArray)

#for gene selection, only use genes (not repetitive elements)


sel_cts <- dds_cts[,diagnosic_1031$Sample]
obj <- seqGlue::calc_dispersion(as.matrix(sel_cts), removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance
ubtf_kmt2a_genes <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(ubtf_kmt2a_genes) #6442


in_vst <- SummarizedExperiment::assay(vst)[ubtf_kmt2a_genes,diagnosic_1031$Sample]

dim(in_vst)
head(in_vst[,1:5]) #
```

```{r}
ubtf_kmt2a_cols <- c("FLT3.UBTF.KMT2A_Groups", "AML_Subtype", "Batch", "Tissue")

# sapply(diagnostic_samples[,ubtf_kmt2a_cols], function(x) length(unique(x)))
sapply(diagnostic_samples[,ubtf_kmt2a_cols], function(x) table(x))

  
ubtf_kmt2a_cc <- list(get_palette("npg",11),
                      get_palette("Accent", 11), 
                      get_palette("simpsons",3),
                      get_palette("rickandmorty", 4)) 


ubtf_kmt2a_cc <- lapply(1:length(ubtf_kmt2a_cc), function(i) set_names(ubtf_kmt2a_cc[[i]],
                                                      unique(diagnostic_samples[,ubtf_kmt2a_cols[i]]))) %>% 
    set_names(ubtf_kmt2a_cols)
ubtf_kmt2a_cc <- sapply(ubtf_kmt2a_cc, function(x){x[grep("All others|No.Primary.Fusion", names(x))] <- "azure2"; return(x)})
ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups["FLT3.ITD+/KMT2A.PTD+"] <- "darkorchid1"
ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups["FLT3.ITD+/UBTF.ITD+"] <- "dodgerblue"
ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups["KMT2A.PTD+"] <- "darkorchid4"
ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups["UBTF.ITD+"] <- "dodgerblue4"


# ubtf_kmt2a_cc



par(mar=c(15,5,5,5))
input <- ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups
barplot(rep(1,length(input)), col=input, names.arg = names(input), las=2)
```

```{r}
umap_bulk_kmt2a <- UMAP_workflow(TFIDF_Matrix = in_vst, 
                           scale_data=FALSE,
                           input_features = ubtf_kmt2a_genes,
                            samples_vector = diagnosic_1031$Sample,
                            sample_info_df = diagnosic_1031,
                            Columns_for_Plots = ubtf_kmt2a_cols,
                            cc = ubtf_kmt2a_cc, 
                            addl_color_vector=rainbow(n=40),
                            min_dist = 0.01,
                            n_neighbors=15,
                            k2=10,
                            res2=0.025)

umap_bulk_kmt2a_df <- umap_bulk_kmt2a$umap_res
# dir.create("KMT2A.PTD_UBTF.ITD_UMAP")
# saveRDS(umap_bulk_kmt2a, "KMT2A.PTD_UBTF.ITD_UMAP/TARGET_AML_KMT2A.PTD_UBTF.ITD_minDist0.01_nn15_FullCohort_UMAP.RDS")
```

```{r}
check <- c(817447, 871386)
check_usi <- orig %>% 
  filter(Reg. %in% check) %>% 
  select(USI, Reg., Protocol)

# umap_bulk_kmt2a_df %>%
#   filter(USI1 %in% check_usi$USI) %>%
#   select(Sample, matches("cluster"),FLT3.UBTF.KMT2A_Groups) #both in cluster 3 (or 13)

# table(umap_bulk_kmt2a_df$FLT3.UBTF.KMT2A_Groups)
# 33 FLT3.ITD+/UBTF.ITD+
# 11 UBTF.ITD+

# umap_bulk_kmt2a_df %>%
#   filter(USI1 %in% check_usi$USI | cluster_k12=="3") %>%
#   select(Sample, matches("cluster"), FLT3.UBTF.KMT2A_Groups) %>%
#   group_by(FLT3.UBTF.KMT2A_Groups) %>%
#   dplyr::count()
```

```{r fig.height=15, fig.width=12}
# umap_bulk_kmt2a$umap_2D_scatter
outlier_check <- ggplot() +
  geom_point(data=filter(umap_bulk_kmt2a_df,grepl("FLT3.ITD\\-|NBM|CD34_PB", FLT3.UBTF.KMT2A_Groups)),
       aes(x=x, y=y, color=FLT3.UBTF.KMT2A_Groups), size=1, alpha=0.4) +
  geom_point(data=filter(umap_bulk_kmt2a_df,grepl("FLT3.ITD\\+|KMT2A.PTD\\+|UBTF.ITD+", FLT3.UBTF.KMT2A_Groups)),
       aes(x=x, y=y, color=FLT3.UBTF.KMT2A_Groups), size=1.2, alpha=0.65) +

  ggrepel::geom_text_repel(data=filter(umap_bulk_kmt2a_df,USI1 %in% check_usi$USI), 
                           aes(x=x, y=y, label=USI1),
                           segment.size=0.25, 
                           min.segment.length=unit(0.01,"mm"), 
                           nudge_x = c(1.75)) +
  guides(color=guide_legend(nrow=4,byrow=TRUE, override.aes=list(size=3))) +
  scale_color_manual(values=ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups) +
  theme_classic() +
  theme(legend.position = "top",
        legend.title = element_blank(),
        plot.margin = margin(r=3,unit = "mm"))
  
 # outlier_check 

# pdf("KMT2A.PTD_UBTF.ITD_UMAP/TARGET_AML_FLT3.ITD_UBTF.ITD_KMT2A.PTD_Leiden_Clusters_UMAP.pdf", height = 15, width = 12)
# (outlier_check + umap_bulk_kmt2a$cluster_plots1$scatter) / umap_bulk_kmt2a$cluster_plots1$barplot
# dev.off()
```

```{r}
info_cols <- c("Sample",ubtf_kmt2a_cols[-c(3:4)],
               c("Primary.Fusion","Mutations.Category","EFS.event.type.ID"))

# all(info_cols %in% colnames(umap_bulk_kmt2a$umap_res))

plotly_kmt2a <- scatter_plots_3d(umap_workflow_res = umap_bulk_kmt2a, 
                 Group_Column = "FLT3.UBTF.KMT2A_Groups", 
                 blackbg = FALSE,
                 Cols=info_cols, 
                 ptsize = 4,
                 cc=ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups)

# plotly_kmt2a

# htmlwidgets::saveWidget(plotly_kmt2a, file="TARGET_AML_FLT3.ITD_UBTF.ITD_KMT2A.PTD_and_Fusions_UMAP.html", selfcontained = TRUE)
```


### Diagnostic AAML1031: Primary Fusions and KMT2A-PTD Focus

```{r}
diagnosic_KMT2A.PTD_1031 <- diagnostic_samples %>% 
  filter(grepl("AAML1031",Protocol) | grepl("NBM", Group)) %>% 
  filter(!grepl("All others|^UBTF.ITD+", FLT3.UBTF.KMT2A_Groups)) %>% 
  mutate(FLT3.KMT2A_Groups=case_when(
    grepl("FLT3.+UBTF", FLT3.UBTF.KMT2A_Groups) ~ "FLT3.ITD+",
    TRUE ~ FLT3.UBTF.KMT2A_Groups)) %>% 
  set_rownames(.$Sample)



# table(diagnosic_KMT2A.PTD_1031$Protocol)
# table(diagnosic_KMT2A.PTD_1031$FLT3.KMT2A_Groups) %>%
#   write.csv(.,"KMT2A.PTD_to_Include.csv")
dim(diagnosic_KMT2A.PTD_1031)
```

```{r}
library(DelayedArray)

#for gene selection, only use genes (not repetitive elements)


sel_cts <- dds_cts[,diagnosic_KMT2A.PTD_1031$Sample]
obj <- seqGlue::calc_dispersion(as.matrix(sel_cts), removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance
kmt2a_genes <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(kmt2a_genes) #6962


in_vst <- SummarizedExperiment::assay(vst)[kmt2a_genes,diagnosic_KMT2A.PTD_1031$Sample]


dim(in_vst)
```

```{r}
kmt2a_cols <- c("FLT3.KMT2A_Groups", "AML_Subtype", "Batch", "Tissue")


# sapply(diagnosic_KMT2A.PTD_1031[,kmt2a_cols], function(x) length(table(x)))
# sapply(diagnosic_KMT2A.PTD_1031[,kmt2a_cols], function(x) table(x))

kmt2a_cc <- list(get_palette("jco",7),
                      get_palette("Accent", 8), 
                      get_palette("simpsons",3),
                      get_palette("rickandmorty", 3)) 


kmt2a_cc <- lapply(1:4, function(i) kmt2a_cc[[i]] %>%  set_names(., unique(diagnosic_KMT2A.PTD_1031[,kmt2a_cols[i]]))) %>% 
  set_names(kmt2a_cols)

kmt2a_cc <- sapply(kmt2a_cc, function(x){x[grep("All others|NBM", names(x))] <- "lightblue1"; return(x)})
kmt2a_cc$FLT3.KMT2A_Groups["FLT3.ITD+"] <- "dodgerblue1"



par(mar=c(15,5,5,5))
input <- kmt2a_cc$AML_Subtype
barplot(rep(1,length(input)), col=input, names.arg = names(input), las=2)
```

n_neighbors	
The size of local neighborhood (in terms of number of neighboring sample points) used for manifold approximation. Larger values result in more global views of the manifold, while smaller values result in more local data being preserved.

min_dist	
The effective minimum distance between embedded points. Smaller values will result in a more clustered/clumped embedding where nearby points on the manifold are drawn closer together, while larger values will result on a more even dispersal of points.

```{r message=FALSE}
umap_bulk_kmt2a.ptd  <- UMAP_workflow(TFIDF_Matrix = in_vst, 
                           scale_data=FALSE,
                           input_features = kmt2a_genes,
                            samples_vector = diagnosic_KMT2A.PTD_1031$Sample,
                            sample_info_df = diagnosic_KMT2A.PTD_1031,
                            Columns_for_Plots = kmt2a_cols,
                            cc = kmt2a_cc, 
                            addl_color_vector=rainbow(n=40),
                            min_dist = 0.01,
                            n_neighbors=11,
                            k2=10,
                            res2=0.025)



umap_bulk_kmt2a.ptd_df <- umap_bulk_kmt2a.ptd$umap_res
# saveRDS(umap_bulk_kmt2a.ptd, "KMT2A.PTD_UBTF.ITD_UMAP/TARGET_AML_KMT2A.PTD_minDist0.01_nn11_MajorFusionsOnly_UMAP.RDS")
```

```{r fig.height=6, fig.width=15}
# umap_bulk_kmt2a.ptd$umap_2D_scatter
# umap_bulk_kmt2a.ptd$cluster_plots1

dim_12 <- ggplot() +
  geom_point(data=filter(umap_bulk_kmt2a.ptd_df, !grepl("PTD", FLT3.KMT2A_Groups)), 
            mapping=aes(x=x,y=y, color=FLT3.KMT2A_Groups), alpha=0.5) +
  geom_point(data=filter(umap_bulk_kmt2a.ptd_df, grepl("PTD", FLT3.KMT2A_Groups)), 
            mapping=aes(x=x,y=y, color=FLT3.KMT2A_Groups), 
            size=2) +
  labs(x="UMAP 1", y="UMAP 2") +
  scale_color_manual(values=kmt2a_cc$FLT3.KMT2A_Groups) +
  guides(color=guide_legend(override.aes = list(size=4))) +
  theme_classic() +
  theme(legend.position = "left")


dim_13 <- ggplot() +
  geom_point(data=filter(umap_bulk_kmt2a.ptd_df, !grepl("PTD", FLT3.KMT2A_Groups)), 
            mapping=aes(x=x,y=z, color=FLT3.KMT2A_Groups), alpha=0.5) +
  geom_point(data=filter(umap_bulk_kmt2a.ptd_df, grepl("PTD", FLT3.KMT2A_Groups)), 
            mapping=aes(x=x,y=z, color=FLT3.KMT2A_Groups), 
            size=2, alpha=1.0) +
  labs(x="UMAP 1", y="UMAP 3") +
  scale_color_manual(values=kmt2a_cc$FLT3.KMT2A_Groups) +
  guides(color=guide_legend(override.aes = list(size=4))) +
  theme_classic() +
  theme(legend.position = "none") 


# pdf("TARGET_AML_KMT2A.PTD_minDist0.01_nn11_MajorFusionsOnly_UMAP_2Dscatter.pdf", height = 6, width = 15)
# dim_12 + dim_13
# dev.off()
```

```{r fig.height=7, fig.width=7}
colors_3d_other <- umap_bulk_kmt2a_df %>% 
  filter(!grepl("FLT3.ITD\\+|KMT2A.PTD\\+|UBTF.ITD+", FLT3.UBTF.KMT2A_Groups)) %>% 
  left_join(., data.frame(color=ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups) %>% 
              rownames_to_column("FLT3.UBTF.KMT2A_Groups"),
            by="FLT3.UBTF.KMT2A_Groups") %>% 
  mutate_at(vars(color), ~sapply(., make_alpha, percent=70)) %>% 
  arrange(FLT3.UBTF.KMT2A_Groups)

n1 <- length(as.numeric(as.factor(colors_3d_other$FLT3.UBTF.KMT2A_Groups)))


colors_3d_kmt2a <- umap_bulk_kmt2a_df %>% 
  filter(grepl("FLT3.ITD\\+|KMT2A.PTD\\+|UBTF.ITD+", FLT3.UBTF.KMT2A_Groups)) %>% 
  left_join(., data.frame(color=ubtf_kmt2a_cc$FLT3.UBTF.KMT2A_Groups) %>% 
              rownames_to_column("FLT3.UBTF.KMT2A_Groups"),
            by="FLT3.UBTF.KMT2A_Groups") %>% 
   mutate_at(vars(color), ~sapply(., make_alpha, percent=40)) %>% 
  arrange(FLT3.UBTF.KMT2A_Groups)

n2 <- length(as.numeric(as.factor(colors_3d_kmt2a$colors_3d_kmt2a)))


table(colors_3d_other$FLT3.UBTF.KMT2A_Groups)
table(colors_3d_kmt2a$FLT3.UBTF.KMT2A_Groups)
```

```{r fig.height=10, fig.width=10}
# t <- 200
# p <- 60

# pdf("Figures/TARGET_AML_ETS_Family_Fusions_plot3D_scatter_3D_long.pdf", height = 10, width = 12)
plot3D::scatter3D(x=colors_3d_other$x, y=colors_3d_other$y, z=colors_3d_other$z,
                   bty = "b2",
                  colvar=as.numeric(as.factor(colors_3d_other$FLT3.UBTF.KMT2A_Groups)),
                  col=unique(colors_3d_other$color),
                  pch=19, colkey=FALSE,
                  # theta=t,phi=p, 
                  cex=1.2,
                  scale = FALSE)

plot3D::scatter3D(x=colors_3d_kmt2a$x, y=colors_3d_kmt2a$y, z=colors_3d_kmt2a$z,
                   bty = "b2",
                  colvar=as.numeric(as.factor(colors_3d_kmt2a$FLT3.UBTF.KMT2A_Groups)),
                  col=unique(colors_3d_kmt2a$color),
                  pch=19, colkey=FALSE,
                  # theta=t,phi=p, 
                  cex=1.0,  
                  scale = FALSE,
                  add = TRUE)
# dev.off()
```



```{r}
# info_cols <- c("Sample",Cols, c("Primary.Fusion","Mutations.Category","EFS.event.type.ID"))
# scatter_plots_3d(umap_workflow_res = )
```




## All FLT3-ITD Focus

```{r}
suppressPackageStartupMessages(library(DelayedArray))

cts.filtered.ids <- cts_grch38_ids %>% 
  filter(gene_name %in% rownames(cts.filtered)) 
  
identical(rownames(CPM), cts.filtered.ids$gene_name) #TRUE
genes <- filter(cts.filtered.ids, grepl("ENSG", gene_id)) %>% 
  pull(gene_name)

# Mean vs Dispersion Feature Selection 
sel_counts <-  cts.filtered[genes, FLT3.ITD_only$Sample] 
sel_counts <- sel_counts[rowSums(sel_counts) > 0, ]

obj <- seqGlue::calc_dispersion(obj = as.matrix(sel_counts), 
                                removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance


sg_SF_all <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(sg_SF_all)  #8482 


sg_SF_top1000 <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=1000))
sg_SF_top5000 <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=5000))
```


### Heatmap 

```{r}
gene_lists <- list(sg_SF_all, sg_SF_top5000, sg_SF_top1000)

heatmaps <- lapply(gene_lists, function(x){
  
  p <- FLT3.ITD_only$FLT3.Groups_PoorRisk %>%
  set_names(FLT3.ITD_only$Sample)

  dends_res <- dge_dendrograms(expnData = logCPM,
                  pheno = p,
                  log=TRUE,
                  method = "ward.D2",
                  percent=0.01,
                  add.count=1,
                  createDGE = FALSE,
                  filterTopGenes = FALSE,
                  genelist = x)

  print(table(x %in% rownames(dends_res$TMMCPM)))
  
  anno <- DeGSEA::create_HA_Labs_Hmap(expn=dends_res$TMMCPM,
                                      geneList = x,
                                      cc = cc_heatmap, #Need a test for all names being > length(0)
                                      CDE = FLT3.ITD_only,
                                      cols = names(cc_heatmap))

  ComplexHmap(mat = dends_res$TMMCPM,
              name = "Z-Scores",
              scale=TRUE,
              dge_dendrograms.res = dends_res,
              hmap_anno_obj = anno$annoColumn)

})

# length(heatmaps)
names(heatmaps) <- ls(pattern = "sg_SF")
```


```{r fig.height=10, fig.width=20}
# lapply(names(heatmaps), function(x){
#   pdf(paste0("TARGET_AML_FLT3-ITD_", x,"_TMMCPM_Heatmap_11.05.2021.pdf"), height = 10, width = 20)
#   print(heatmaps[[x]])
#   dev.off()
# })
```


### UMAP

n_neighbors	
The size of local neighborhood (in terms of number of neighboring sample points) used for manifold approximation. Larger values result in more global views of the manifold, while smaller values result in more local data being preserved.

min_dist	
The effective minimum distance between embedded points. Smaller values will result in a more clustered/clumped embedding where nearby points on the manifold are drawn closer together, while larger values will result on a more even dispersal of points. 


```{r message=FALSE}
FLT3_UMAP  <- UMAP_workflow(TFIDF_Matrix = assay(vst)[sg_SF_all, FLT3.ITD_only$Sample], 
                           scale_data=FALSE,
                           input_features = sg_SF_all,
                            samples_vector = FLT3.ITD_only$Sample,
                            sample_info_df = FLT3.ITD_only,
                            Columns_for_Plots = names(cc_heatmap),
                            cc = cc_heatmap, 
                            addl_color_vector=rainbow(n=40),
                            min_dist = 1e-8,
                            n_neighbors=4,
                            k2=10,
                            res2=0.025)


# saveRDS(FLT3_UMAP, "FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_minDistXX_nnXX_UMAP.RDS")
```

```{r fig.height=10, fi.width=10}
FLT3_UMAP$umap_2D_scatter
```



## Low Risk FLT3-ITD

```{r}
FLT3.ITD_LR <- diagnostic_samples %>% 
  filter(!grepl("High risk", FLT3.Groups_FavRisk)) %>%
  filter(grepl("CBFB-MYH11|RUNX1-RUNX1T1", AML_Subtype) | 
           CEBPA.mutation.=="Yes" | NPM.mutation. == "Yes") %>% 
  mutate_at(vars(FLT3.Groups_FavRisk), ~case_when(
    grepl("CBFB-MYH11|RUNX1-RUNX1T1", AML_Subtype) & .=="FLT3-ITD-" ~ paste(AML_Subtype,., sep="/"),
    CEBPA.mutation. == "Yes" & .=="FLT3-ITD-" ~ paste("CEBPA+",., sep="/"),
    NPM.mutation.=="Yes" & .=="FLT3-ITD-" ~ paste("NPM1+", ., sep="/"), 
    grepl("FLT3.ITD\\+\\/NPM\\+$", FLT3.Groups) ~ "NPM1+/FLT3-ITD+",
    grepl("FLT3.ITD\\+\\/CEBPA\\+$", FLT3.Groups) ~ "CEBPA+/FLT3-ITD+",
    TRUE ~ .)) %>%
  set_rownames(.$Sample)


dim(FLT3.ITD_LR)
table(FLT3.ITD_LR$FLT3.Groups_FavRisk)
# length(table(FLT3.ITD_LR$FLT3.Groups_FavRisk)) #7
#table(FLT3.ITD_LR$CEBPA.mutation., FLT3.ITD_LR$NPM.mutation.) #0 positive for both
```

```{r}
table(FLT3.ITD_LR$Primary.Fusion)
table(FLT3.ITD_LR$NPM.mutation.)
table(FLT3.ITD_LR$CEBPA.mutation.)
```

```{r}
suppressPackageStartupMessages(library(DelayedArray))

cts.filtered.ids <- cts_grch38_ids %>% 
  filter(gene_name %in% rownames(cts.filtered)) 
  
identical(rownames(CPM), cts.filtered.ids$gene_name) #TRUE
genes <- filter(cts.filtered.ids, grepl("ENSG", gene_id)) %>% 
  pull(gene_name)

# Mean vs Dispersion Feature Selection 
LR_counts <-  cts.filtered[genes, FLT3.ITD_LR$Sample] 
LR_counts <- LR_counts[rowSums(LR_counts) > 0, ]

obj <- seqGlue::calc_dispersion(as.matrix(LR_counts), removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance

sg_LR_all <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(sg_LR_all)  # 7996

sg_LR_top1000 <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=1000))

sg_LR_top2000 <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=2000))
```

```{r}
cc_LR <- list()
lr_pal <- get_palette("Paired", 8)
cc_LR[["FLT3.Groups_FavRisk"]] <- c(
  "CBFB-MYH11/FLT3.ITD+" = "firebrick4",
  "CBFB-MYH11/FLT3-ITD-" = "firebrick1",
  
  "CEBPA+/FLT3-ITD-" = "#A6CEE3",
  "CEBPA+/FLT3-ITD+" = "#1F78B4", 
  
  "NPM1+/FLT3-ITD-" = "slateblue1", 
  "NPM1+/FLT3-ITD+" = "slateblue4",
  
  "RUNX1-RUNX1T1/FLT3-ITD-" = "burlywood2", 
  "RUNX1-RUNX1T1/FLT3.ITD+" = "burlywood4"
)


par(mar=c(15,5,5,5))
input <- cc_LR$FLT3.Groups_FavRisk
barplot(rep(1,length(input)), col=input, names.arg = names(input), las=2)
```


#### UMAP 

```{r message=FALSE}
FLT3_LR_UMAP  <- UMAP_workflow(TFIDF_Matrix = assay(vst)[sg_LR_all,FLT3.ITD_LR$Sample], 
                           scale_data=FALSE,
                           input_features = sg_LR_all,
                            samples_vector = FLT3.ITD_LR$Sample,
                            sample_info_df = FLT3.ITD_LR,
                            Columns_for_Plots = names(cc_LR),
                            cc = cc_LR, 
                            addl_color_vector=rainbow(n=40),
                            min_dist = 0.01,
                            n_neighbors=11,
                            k2=10,
                            res2=0.025)

# dir.create("FLT3_Cooperating_Muts_Tarlock")
# saveRDS(FLT3_LR_UMAP, "FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_Low_Risk_minDist0.01_nn11_UMAP.RDS")
```

```{r fig.height=10, fig.width=10}
FLT3_LR_UMAP <- readRDS("FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_Low_Risk_minDist0.01_nn11_UMAP.RDS")
# FLT3_LR_UMAP$umap_2D_scatter
```

```{r}
#Clean dataframe for 3D scatter plot
FLT3_LR_UMAP_df <- FLT3_LR_UMAP$umap_res %>% 
  full_join(., data.frame(color=cc_LR$FLT3.Groups_FavRisk,
                          FLT3.Groups_FavRisk=names(cc_LR$FLT3.Groups_FavRisk)),
            by="FLT3.Groups_FavRisk") %>% 
  rowwise() %>%
  mutate_at(vars(color), ~case_when(
    grepl("FLT3.ITD-", FLT3.Groups_FavRisk) ~ make_alpha(color, percent=80),
    grepl("FLT3.ITD\\+", FLT3.Groups_FavRisk) ~ make_alpha(color, percent=20)
  )) %>%
  ungroup() %>%
  mutate(Major_Subtype=case_when(
    grepl("RUNX1|CBFB", FLT3.Groups_FavRisk) ~ Primary.Fusion,
    grepl("NPM1", FLT3.Groups_FavRisk) ~ "NPM1", 
    grepl("CEBPA", FLT3.Groups_FavRisk) ~ "CEBPA")) %>% 
  arrange(FLT3.Groups_FavRisk)



dim(FLT3_LR_UMAP_df)
# table(FLT3_LR_UMAP_df$color,
#       as.numeric(as.factor(FLT3_LR_UMAP_df$FLT3.Groups_FavRisk)))
```

```{r}
# table(FLT3_LR_UMAP_df$FLT3.Groups_FavRisk, FLT3_LR_UMAP_df$cluster_k4)
# table(FLT3_LR_UMAP_df$Major_Subtype, FLT3_LR_UMAP_df$FLT3.Groups_FavRisk)
  FLT3_LR_UMAP$cluster_plots1
calc_cluster_purity(FLT3_LR_UMAP_df$Major_Subtype, FLT3_LR_UMAP_df$cluster_k4) %>% 
  round(., digits = 3)
```

```{r fig.height=10, fig.width=12}
group_1 <- FLT3_LR_UMAP_df %>% 
  filter(grepl("FLT3.ITD-", FLT3.Groups_FavRisk))

group_2 <- FLT3_LR_UMAP_df %>% 
  filter(grepl("FLT3.ITD\\+", FLT3.Groups_FavRisk))

t <- 40
p <- 15


# pdf("Figures/TARGET_AML_FLT3.ITD_LowRisk_plot3D_scatter_3D_long.pdf", height = 10, width = 12)
par(mar=c(3,5,18,5))
plot3D::scatter3D(x=group_1$x, y=group_1$y, z=group_1$z,
                   bty = "b2",
                  colvar=as.numeric(as.factor(group_1$FLT3.Groups_FavRisk)),
                  col=unique(group_1$color),
                  pch=19, 
                  colkey=list(side=2, plot=TRUE,
                              length=0.25,
                              labels=unique(group_1$FLT3.Groups_FavRisk),
                              at=c(1:4),
                              dist=-0.4,
                              shift=0.375,
                              cex.axis=1.0,
                              addlines=TRUE),
                  theta=t,phi=p,
                  cex=0.85,
                  add=FALSE,
                  scale = FALSE)

plot3D::scatter3D(x=group_2$x, y=group_2$y, z=group_2$z,
                   bty = "b2",
                  colvar=as.numeric(as.factor(group_2$FLT3.Groups_FavRisk)),
                  col=unique(group_2$color),
                  pch=19, 
                  colkey=list(side=4, plot=TRUE,
                              length=0.25, 
                              tick=TRUE,
                              labels=unique(group_2$FLT3.Groups_FavRisk),
                              at=c(1:4),
                              dist=-0.25, 
                              shift=0.375,
                              cex.axis=1.0,
                              addlines=TRUE),
                  theta=t,phi=p,
                  cex=1.1,
                  add=TRUE,
                  scale = FALSE)

lim <-length(unique(FLT3_LR_UMAP_df$color))
# dev.off()
```


### Heatmap

```{r}
gene_lists <- list(sg_TMMCPM_LR_all, sg_TMMCPM_LR_top2000, sg_TMMCPM_LR_top1000)

heatmaps_LR <- lapply(gene_lists, function(x){
  
  #define the phenotypes of each sample
  p <- FLT3.ITD_LR$FLT3.Groups_FavRisk %>%
  set_names(FLT3.ITD_LR$Sample)
  
  dends_res <- dge_dendrograms(expnData = logCPM,
                  pheno = p,
                  log=TRUE,
                  method = "ward.D2",
                  percent=0.01,
                  add.count=1,
                  createDGE = FALSE,
                  filterTopGenes = FALSE,
                  genelist = x)

  print(table(x %in% rownames(dends_res$TMMCPM)))

  #update the color codes
  codes <- cc_heatmap[-grep("FLT3.Groups_PoorRisk", names(cc_heatmap))]
  codes <- sapply(names(codes), function(x){
        idx <- names(codes[[x]]) %in% unique(FLT3.ITD_LR[[x]])
        codes[[x]][idx]
  })

  anno <- DeGSEA::create_HA_Labs_Hmap(expn=dends_res$TMMCPM,
                                      geneList = x,
                                      cc = codes, #Need a test for all names being > length(0)
                                      CDE = FLT3.ITD_LR,
                                      cols = names(codes))
  
  ComplexHmap(mat = dends_res$TMMCPM,
              name = "Z-Scores",
              scale=TRUE,
              dge_dendrograms.res = dends_res,
              hmap_anno_obj = anno$annoColumn)

})


names(heatmaps_LR) <- ls(pattern = "sg_TMMCPM_LR")
```

```{r}
lapply(names(heatmaps_LR), function(x){
  pdf(paste0("Figures/TARGET_AML_FLT3-ITD_", x,"_Heatmap_10.18.2021.pdf"), height = 10, width = 20)
  print(heatmaps_LR[[x]])
  dev.off()})
```

## High Risk FLT3-ITD

```{r}
FLT3_HR <- diagnostic_samples %>% 
  filter(FLT3.ITD_Groups_Simple=="High risk/FLT3-ITD+" | grepl("NUP98-NSD1", Primary.Fusion)) %>% 
  mutate_at(vars(FLT3.Groups_PoorRisk), ~ifelse(FLT3.ITD.positive.=="No", Primary.Fusion, .)) %>% 
  mutate(FLT3.Groups_PoorRisk=as.factor(FLT3.Groups_PoorRisk)) %>% 
  set_rownames(.$Sample)

dim(FLT3_HR)
# table(FLT3_HR$Primary.Fusion, FLT3_HR$FLT3.Groups_PoorRisk)
# table(FLT3_HR$FLT3.ITD_Groups_Simple, useNA = 'always')
# table(FLT3_HR$FLT3.Groups_PoorRisk)
```


```{r}
cc_HR <- cc_heatmap["FLT3.Groups_PoorRisk"]
cc_HR["FLT3.Groups_PoorRisk"] <- lapply(names(cc_HR), function(x) cc_HR[[x]][intersect(names(cc_HR[[x]]), FLT3_HR[[x]])])
cc_HR[["FLT3.Groups_PoorRisk"]]["NUP98-NSD1"] <- "darkgrey" 
cc_HR[["FLT3.Groups_PoorRisk"]]["DEK-NUP214/FLT3.ITD+"] <- "#EFC000FF"


cc_HR[["AML_Subtype"]] <- brewer.pal(length(unique(FLT3_HR$AML_Subtype)),"Paired") %>% 
  set_names(unique(FLT3_HR$AML_Subtype))
  
# cc_HR[["AML_Subtype"]] <- brewer.pal(length(unique(FLT3_HR$AML_Subtype)),"Set1") %>%
#   set_names(unique(FLT3_HR$AML_Subtype))
# cc_HR
par(mar=c(12s,5,5,5))
input <- cc_HR$FLT3.Groups_PoorRisk
barplot(rep(1, length(input)), col=input, names.arg = names(input), las=2)
```


### UMAP

```{r}
suppressPackageStartupMessages(library(DelayedArray))


cts.filtered.ids <- cts_grch38_ids %>% 
  filter(gene_name %in% rownames(cts.filtered)) 
  
identical(rownames(CPM), cts.filtered.ids$gene_name) #TRUE
genes <- filter(cts.filtered.ids, grepl("ENSG", gene_id)) %>% 
  pull(gene_name)

# Mean vs Dispersion Feature Selection 
sel_norm_counts <-  dds_cts[genes,FLT3_HR$Sample] 
dim(sel_norm_counts)

obj <- seqGlue::calc_dispersion(as.matrix(sel_norm_counts), removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance

sg_SF_HR_all <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(sg_SF_HR_all)  # 8081
```


k	
Integer number of nearest neighbors to use when creating the k nearest neighbor graph for Louvain/Leiden clustering. k is related to the resolution of the clustering result, a bigger k will result in lower resolution and vice versa. Default is 20.

min_dist	
The effective minimum distance between embedded points. Smaller values will result in a more clustered/clumped embedding where nearby points on the manifold are drawn closer together, while larger values will result on a more even dispersal of points. The value should be set relative to the spread value, which determines the scale at which embedded points will be spread out.

n_neighbors	
The size of local neighborhood (in terms of number of neighboring sample points) used for manifold approximation. Larger values result in more global views of the manifold, while smaller values result in more local data being preserved. In general values should be in the range 2 to 100.

### Standard Method 

```{r message=FALSE}
FLT3_HR_UMAP  <- UMAP_workflow(TFIDF_Matrix = assay(vst)[sg_SF_HR_all, FLT3_HR$Sample],
                           scale_data=FALSE,
                           input_features = sg_SF_HR_all,
                            samples_vector = FLT3_HR$Sample,
                            sample_info_df = FLT3_HR,
                            Columns_for_Plots = names(cc_HR),
                            cc = cc_HR,
                            addl_color_vector=rainbow(n=40),
                            min_dist = 1e-3,
                            n_neighbors=3,
                            k2=15,
                            res2=0.01) 

# saveRDS(FLT3_HR_UMAP, "FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_High_Risk_minDist.1e-8_nn3_UMAP.RDS")
# saveRDS(FLT3_HR_UMAP, "FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_High_Risk_minDist0.1_nn3_UMAP.RDS")

FLT3_HR_UMAP <- readRDS("FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_High_Risk_minDist0.1_nn3_UMAP.RDS")
```

```{r}
#for 3D cluster plots
FLT3_HR_UMAP.df <- FLT3_HR_UMAP$umap_res %>% 
  select(x,y,z,Sample, FLT3.Groups_PoorRisk, AML_Subtype, Primary.Fusion, cluster_k2) %>% 
  # mutate_at(vars(FLT3.Groups_PoorRisk), ~case_when(
  #   AML_Subtype=="NUP98-NSD1" ~ "NUP98-NSD1/FLT3-ITD+",
  #   TRUE ~ .)) %>% 
  left_join(., data.frame(colors=cc_HR[["FLT3.Groups_PoorRisk"]], 
                          FLT3.Groups_PoorRisk=names(cc_HR[["FLT3.Groups_PoorRisk"]])),
            by="FLT3.Groups_PoorRisk") %>% 
  rowwise() %>% 
  mutate_at(vars(colors), ~make_alpha(.,percent=30)) %>% 
  ungroup() %>% 
  mutate(NUP98.NSD1=case_when(
    grepl("NUP98-NSD1", Primary.Fusion) ~ Primary.Fusion,
    TRUE ~ "otherPRM"))


table(FLT3_HR_UMAP.df$NUP98.NSD1)
```

```{r fig.height=12, fig.width=15}
# pdf("TARGET_AML_FLT3.ITD_High_Risk_minDist0.1_nn3_2D_scatter_cluster.pdf", height = 12.5, width = 15)
(FLT3_HR_UMAP$umap_2D_scatter$FLT3.Groups_PoorRisk + FLT3_HR_UMAP$umap_2D_scatter$AML_Subtype) /
(FLT3_HR_UMAP$cluster_plots1$scatter + FLT3_HR_UMAP$cluster_plots1$barplot)
# dev.off()
```

```{r}
# calc_cluster_purity(FLT3_HR_UMAP$umap_res$cluster_k2, FLT3_HR_UMAP$umap_res$FLT3.Groups_PoorRisk)

# calc_cluster_purity(FLT3_HR_UMAP.df$cluster_k2, FLT3_HR_UMAP.df$NUP98.NSD1) %>% 
#   round(., digits=2)
```

```{r}
# table(FLT3_HR_UMAP$umap_res$cluster_k2)

confmat_NSD1 <- FLT3_HR_UMAP$umap_res %>% 
  mutate(NSD1=case_when(
    Primary.Fusion == "NUP98-NSD1" ~ Primary.Fusion,
    TRUE ~ "otherPRM")) %>% 
  group_by(NSD1, cluster_k2) %>% 
  dplyr::count() %>% 
  ungroup() %>% 
  pivot_wider(names_from=NSD1, values_from=n) %>% 
  column_to_rownames("cluster_k2")


addmargins(as.matrix(confmat_NSD1))

fisher.test(confmat_NSD1)
```

```{r}
FLT3_HR_3D <- scatter_plots_3d(umap_workflow_res = FLT3_HR_UMAP,
                                      Group_Column = "FLT3.Groups_PoorRisk", 
                                      Cols = c("Sample","FLT3.Groups_PoorRisk", 
                                               "AML_Subtype","Mutations.Category", "Primary.Fusion","EFS.event.type.ID"),
                                      cc = cc_HR$FLT3.Groups_PoorRisk,
                                      blackbg = FALSE)

FLT3_HR_3D
# htmlwidgets::saveWidget(FLT3_HR_3D, file="TARGET_AML_FLT3.ITD_High_Risk_all_NUP98.NSD1_minDist0.5_nn3_pca10_UMAP.html", selfcontained = TRUE)
```


```{r fig.height=10, fig.width=10}
# colors_to_plot <- list()
# colors_to_plot[["FLT3.Groups_PoorRisk"]] <- cc_HR[["FLT3.Groups_PoorRisk"]][-c(3:4)]
# colors_to_plot[["FLT3.Groups_PoorRisk"]]["NUP98-NSD1/FLT3-ITD+"] <- "#4DBBD5FF"
t <- 20
p <- 30


# pdf("Figures/TARGET_AML_FLT3.ITD_HighRisk_triplePositivesColored_plot3D_scatter_3D_12.17.2021.pdf", height = 10, width = 12)
par(mar=c(3,5,18,5))
plot3D::scatter3D(x=FLT3_HR_UMAP.df$x, y=FLT3_HR_UMAP.df$y, z=FLT3_HR_UMAP.df$z,
                   bty = "b2",
                  colvar=as.numeric(as.factor(FLT3_HR_UMAP.df$FLT3.Groups_PoorRisk)),
                  col=unique(FLT3_HR_UMAP.df$colors),
                  pch=19, 
                  colkey=list(side=2, plot=TRUE,
                              length=0.25,
                              labels=unique(FLT3_HR_UMAP.df$FLT3.Groups_PoorRisk),
                              at=c(1:4),
                              dist=-0.2,
                              shift=0.375,
                              cex.axis=1.0,
                              addlines=TRUE),
                  theta=t,
                  phi=p,
                  cex=1.0,
                  add=FALSE,
                  scale = TRUE)

# dev.off()
# lim <-length(unique(FLT3_LR_UMAP_df$color))
```

### PCA Step Included

```{r}
vst_HR <- DESeq2::vst(dds[, FLT3_HR$Sample], blind = TRUE)
dim(vst_HR)
```

OK params: with VST_HR
scale=FALSE,
min_dist = 1e-2,
pca=10,
n_neighbors = 3


BEST so far: with VST_HR
scale=FALSE,
min_dist = 5e-1,
pca=10,
n_neighbors = 3
                                         
                                         
```{r message=FALSE}
umap_testing <- UMAP_function(data_matrix = assay(vst_HR)[sg_SF_HR_all, ],
                              metadata = FLT3_HR,
                              N=3,
                              scale=FALSE,
                              min_dist = 5e-1,
                              pca=10,
                              n_neighbors = 3)

```

```{r fig.height=10, fig.width=10}
output <- umap_testing$umap_df
  # left_join(.,  FLT3_HR, by="Sample")

# ggplot(output, aes(x=x,y=y, color=FLT3.Groups_PoorRisk)) +
#   geom_point() +
#   scale_color_manual(values = cc_HR$FLT3.Groups_PoorRisk) +
#   theme_classic()


output <- list("umap_res"=output)

# output_3D <- scatter_plots_3d(umap_workflow_res = output,
#                                       Group_Column = "FLT3.Groups_PoorRisk", 
#                                       Cols = c("Sample","FLT3.Groups_PoorRisk", 
#                                                "AML_Subtype","Mutations.Category", "Primary.Fusion","EFS.event.type.ID"),
#                                       cc = cc_HR$FLT3.Groups_PoorRisk,
#                                       blackbg = FALSE)

# output_3D
# htmlwidgets::saveWidget(output_3D, file="TARGET_AML_FLT3.ITD_High_Risk_all_NUP98.NSD1_minDist0.5_nn3_pca10_UMAP.html", selfcontained = TRUE)

FLT3_HR_Test.df <- output$umap_res %>% 
  select(x,y,z,Sample, FLT3.Groups_PoorRisk, AML_Subtype) %>% 
  # mutate_at(vars(FLT3.Groups_PoorRisk), ~case_when(
  #   AML_Subtype=="NUP98-NSD1" ~ "NUP98-NSD1/FLT3-ITD+",
  #   TRUE ~ .)) %>% 
  left_join(., data.frame(colors=cc_HR[["FLT3.Groups_PoorRisk"]], 
                          FLT3.Groups_PoorRisk=names(cc_HR[["FLT3.Groups_PoorRisk"]])),
            by="FLT3.Groups_PoorRisk") %>% 
  rowwise() %>% 
  mutate_at(vars(colors), ~make_alpha(.,percent=30)) %>% 
  ungroup()


t <- 25
p <- 15


# pdf("Figures/TARGET_AML_FLT3.ITD_HighRisk_allNUP98.NSD1_minDist0.5_nn3_pca10_plot3D_scatter_3D_wide.pdf", height = 10, width = 12)
par(mar=c(3,5,18,5))
plot3D::scatter3D(x=FLT3_HR_Test.df$x, y=FLT3_HR_Test.df$y, z=FLT3_HR_Test.df$z,
                   bty = "b2",
                  colvar=as.numeric(as.factor(FLT3_HR_Test.df$FLT3.Groups_PoorRisk)),
                  col=unique(FLT3_HR_Test.df$colors),
                  pch=19, 
                  colkey=list(side=2, plot=TRUE,
                              length=0.25,
                              labels=unique(FLT3_HR_Test.df$FLT3.Groups_PoorRisk),
                              at=c(1:5),
                              dist=-0.2,
                              shift=0.375,
                              cex.axis=1.0,
                              addlines=TRUE),
                  theta=t,
                  phi=p,
                  cex=1.0,
                  add=FALSE,
                  scale = TRUE)
# dev.off()
```




## All FLT3-ITD with Other AML

```{r}
FLT3.ITD_MajorFusions <- diagnostic_samples %>% 
  filter(FLT3.ITD.positive.=="Yes" | grepl("NBM|CBFB-MYH11|RUNX1-RUNX1T1|DEK-NUP214|NUP98-NSD1|KMT2A", AML_Subtype) |
           NPM.mutation.=="Yes" | CEBPA.mutation.=="Yes") %>% 
  mutate_at(vars(FLT3.Groups_PoorRisk), ~case_when(
    .=="FLT3-ITD-" & NPM.mutation.=="Yes" ~ "NPM1+/FLT3.ITD-",
    .=="FLT3-ITD-" & CEBPA.mutation.=="Yes" ~ "CEBPA+/FLT3.ITD-",
    .=="FLT3-ITD-" ~ paste0(AML_Subtype,"/FLT3.ITD-"),
    grepl("FLT3.ITD\\+\\/NPM\\+$", FLT3.Groups) ~ "NPM1+/FLT3.ITD+",
    grepl("FLT3.ITD\\+\\/CEBPA\\+$", FLT3.Groups) ~ "CEBPA+/FLT3.ITD+",
    .=="Low risk/FLT3-ITD+" ~ FLT3.Groups_FavRisk,
    TRUE ~ .)) %>% 
  set_rownames(.$Sample)


# table(FLT3.ITD_MajorFusions$FLT3.Groups_PoorRisk)
# length(unique(FLT3.ITD_MajorFusions$FLT3.Groups_PoorRisk)) #16 groups
```

```{r}
suppressPackageStartupMessages(library(DelayedArray))


cts.filtered.ids <- cts_grch38_ids %>% 
  filter(gene_name %in% rownames(cts.filtered)) 
  
identical(rownames(CPM), cts.filtered.ids$gene_name) #TRUE
genes <- filter(cts.filtered.ids, grepl("ENSG", gene_id)) %>% 
  pull(gene_name)

# Mean vs Dispersion Feature Selection 
sel_norm_counts <-  dds_cts[genes,FLT3.ITD_MajorFusions$Sample] 

obj <- seqGlue::calc_dispersion(as.matrix(sel_norm_counts), removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance

sg_SF_MajorFus_all <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(sg_SF_MajorFus_all)  # 6793
```

```{r}
pal <- get_palette("npg", 10)

cc_heatmap2 <- list()
# cc_heatmap2[["Batch"]] <- viridis(n=length(unique(diagnostic_samples$Batch))) %>% 
#   set_names(unique(diagnostic_samples$Batch))
# 
# cc_heatmap2[["Tissue"]] <- viridis(n=length(unique(unique(diagnostic_samples$Tissue)))) %>% 
#   set_names(unique(diagnostic_samples$Tissue))

cc_heatmap2[["FLT3.Groups_PoorRisk"]] <- c(
  "All others/FLT3.ITD+"="grey50",
  "DEK-NUP214/FLT3.ITD+"="#00A087FF",
  "DEK-NUP214/FLT3.ITD-"="#91D1C2FF",
  
  "FLT3.ITD+/WT1+"="goldenrod1",
  "KMT2A/FLT3.ITD-"="khaki2",

  "NUP98-NSD1/FLT3.ITD+"="#4DBBD5FF",
  "NUP98-NSD1/FLT3.ITD+/WT1+"="#3C5488FF",
  "NUP98-NSD1/FLT3.ITD-"="darkslategray2",
  
  "CBFB-MYH11/FLT3.ITD+" = "firebrick4",
  "CBFB-MYH11/FLT3.ITD-" = "firebrick1",

  "CEBPA+/FLT3.ITD-" = "#A6CEE3",
  "CEBPA+/FLT3.ITD+" = "#1F78B4", 
  
  "NPM1+/FLT3.ITD-" = "slateblue1", 
  "NPM1+/FLT3.ITD+" = "slateblue4",
  
  "RUNX1-RUNX1T1/FLT3.ITD-" = "burlywood2", 
  "RUNX1-RUNX1T1/FLT3.ITD+" = "burlywood4",
 
  "NBM"="black"
  # "FLT3-ITD-"="azure2",
  # "CD34_PB"="grey50"
)

cc_heatmap2[["AML_Subtype"]] <- get_palette("Set1", length(unique(diagnostic_samples$AML_Subtype))) %>% 
  set_names(unique(diagnostic_samples$AML_Subtype))
cc_heatmap2[["AML_Subtype"]]["AML, NOS"] <-  "grey80"
cc_heatmap2[["AML_Subtype"]]["No.Primary.Fusion"] <- "wheat2" 
cc_heatmap2[["AML_Subtype"]]["NUP98-KDM5A"] <- "royalblue1"
cc_heatmap2[["AML_Subtype"]]["CBFB-MYH11"] <- "sienna"
cc_heatmap2[["AML_Subtype"]]["NBM"] <- "black"


par(mar=c(15,5,5,5))
input <- cc_heatmap2$AML_Subtype
barplot(rep(1,length(input)),
        col = input,
        names.arg = names(input),
        las=2)
```

### UMAP 

```{r}
UMAP_FLT3_MajorFusions <- UMAP_workflow(TFIDF_Matrix = assay(vst)[,FLT3.ITD_MajorFusions$Sample], 
                                        input_features = sg_SF_MajorFus_all,
                                        samples_vector = FLT3.ITD_MajorFusions$Sample, 
                                        sample_info_df = FLT3.ITD_MajorFusions,
                                        cc = cc_heatmap2, 
                                        Columns_for_Plots = names(cc_heatmap2),
                                        min_dist = 1e-5,
                                        n_neighbors = 55,
                                        k2 = 15, res=0.01)


# saveRDS(UMAP_FLT3_MajorFusions,"FLT3_Cooperating_Muts_Tarlock/TARGET_AML_FLT3.ITD_and_MajorFusions_minDist.1e-5_nn55_UMAP.RDS")
```

```{r fig.height=10, fig.width=10}
# UMAP_FLT3_MajorFusions$umap_2D_scatter
UMAP_FLT3_MajorFusions$cluster_plots1
```

```{r}
FLT3_MajorFusions <- scatter_plots_3d(umap_workflow_res = UMAP_FLT3_MajorFusions,
                                      Group_Column = "FLT3.Groups_PoorRisk", 
                                      Cols = c("Sample","FLT3.Groups_PoorRisk", "AML_Subtype","Mutations.Category", "Primary.Fusion","EFS.event.type.ID"),
                                      cc = cc_heatmap2$FLT3.Groups_PoorRisk,
                                      blackbg = FALSE)
```

```{r}
# FLT3_MajorFusions
# htmlwidgets::saveWidget(FLT3_MajorFusions, file="TARGET_AML_FLT3.ITD_and_MajorFusions_minDist.1e-5_nn55_UMAP.html", selfcontained = TRUE)

```

### Heatmap

```{r}
gene_lists <- list(sg_TMMCPM_dx_all, sg_TMMCPM_dx_top1000, sg_TMMCPM_dx_top2000)

heatmaps_dx <- lapply(gene_lists, function(x){
  
  p <- diagnostic_samples$FLT3.Groups_PoorRisk %>%
  set_names(diagnostic_samples$Sample)


  dends_res <- dge_dendrograms(expnData = logCPM,
                  pheno = p,
                  log=TRUE,
                  method = "ward.D2",
                  percent=0.01,
                  add.count=1,
                  createDGE = FALSE,
                  filterTopGenes = FALSE,
                  genelist = x)


  # in_vst <- SummarizedExperiment::assay(vst)[x,]
  # 
  # d1 <- dist(t(in_vst), method = "euclidean", diag = FALSE,
  #            upper = FALSE) #sample distances WITHOUT SCALING
  # d2 <- dist(in_vst, method = "euclidean", diag = FALSE,
  #            upper = TRUE) #gene distances WITHOUT SCaling
  # 
  # samp.c1 <- hclust(d1, method = "ward.D2", members = NULL) #sample clustering
  # gene.c2 <- hclust(d2, method = "ward.D2", members = NULL) #gene clustering
  # 
  # dends_res <- list(in_vst,samp.c1,gene.c2)
  # names(dends_res) <- c("TMMCPM","samp.c1", "gene.c2")

  #NEED TO FIX: 1) make USI or sample_ID column into a variable
  # Provide an erorr or warning when the subset dataframe has zero rows.
  
  print(table(x %in% rownames(dends_res$TMMCPM)))
  
  anno <- DeGSEA::create_HA_Labs_Hmap(expn=dends_res$TMMCPM,
                                      geneList = x,
                                      cc = cc_heatmap2,
                                      CDE = diagnostic_samples,
                                      cols = names(cc_heatmap2))


  ComplexHmap(mat = dends_res$TMMCPM, 
              name = "Z-Scores",
              scale=TRUE,
              dge_dendrograms.res = dends_res,
              hmap_anno_obj = anno$annoColumn)

  
  
})

# length(heatmaps)
names(heatmaps_dx) <- ls(pattern = "sg_TMMCPM_dx")
```

```{r fig.height=10, fig.width=20}
# lapply(names(heatmaps_dx), function(x){
#   pdf(paste0("Figures/TARGET_AML_FLT3-ITD_", x,"_Heatmap_10.13.2021.pdf"), height = 10, width = 20)
#   print(heatmaps_dx[[x]])
#   dev.off()})
```


## UMAP

```{r}
adjusted_subset <- ComBat_seq(counts = as.matrix(cts.filtered[sg_all, FLT3.ITD_only$Sample]),
                       batch=FLT3.ITD_only$Batch_noSingletons,
                       group=FLT3.ITD_only$FLT3.Groups_PoorRisk,
                       full_mod = TRUE)

# adjusted_test
range(adjusted_test)
head(adjusted_test[,1:5])
```

```{r}
adjusted <- ComBat_seq(counts = as.matrix(cts.filtered[,diagnostic_samples$Sample]),
                       batch=diagnostic_samples$Batch_noSingletons,
                       group=diagnostic_samples$FLT3.Groups_PoorRisk,
                       full_mod = TRUE)

dim(adjusted)
# saveRDS(adjusted,"TARGET_AML_FLT3.ITD_co-occuring_muts_ComBatseq.RDS")
```

```{r}
cts.filtered.ids <- cts_grch38_ids %>% 
  filter(gene_name %in% rownames(cts.filtered)) %>% 
  filter(grepl("ENSG", gene_id))

# Im wondering if I want to do VST on only FLT3-ITD or the whole dataset to capture variablility of FLT3-ITD in the presence of other AML subtypes?
# or do I want to really look at intra-FLT3-ITD variablility?
dds <- DESeq2::DESeqDataSetFromMatrix(round(cts.filtered[cts.filtered.ids$gene_name, ], digits = 0),
                                      colData = diagnostic_samples,
                                      design = ~ 1)

# normFactors <- normFactors / exp(rowMeans(log(normFactors)))
# normalizationFactors(dds) <- normFactors
dds <- estimateSizeFactors(dds)
dds <- estimateDispersions(dds)
dds_cts <- counts(dds, normalized=TRUE)

# plotDispEsts(dds)
vst <- DESeq2::vst(dds, blind = TRUE)
dim(vst)
```

with the vst as input:
the condition has length > 1 and only the first element will be usedRemoving 8 outliers
no observations informative at iteration 1glm.fit: algorithm did not convergeError in glm.fit(x = numeric(0), y = numeric(0), weights = NULL, start = c(1e-06,  : object 'fit' not found


with cpm as input:
the condition has length > 1 and only the first element will be usedRemoving 211 outliers
NaNs producedstep size truncated due to divergenceNaNs produced  

```{r}
library(DelayedArray)

#for gene selection, only use genes (not repetitive elements), and only select those from the populations of interest FLT3-ITD
# sel_cts <- cts.filtered[cts.filtered.ids$gene_name, FLT3.ITD_only$Sample]
# sel_cts <- SummarizedExperiment::assay(vst)
sel_cts <- dds_cts[,FLT3.ITD_only$Sample]
obj <- seqGlue::calc_dispersion(as.matrix(sel_cts), removeOutliers = TRUE) #removes outlier genes/transcripts based on cooks distance
sg_all <- seqGlue::get_selected_genes(seqGlue::select_genes(obj, top_n=NULL))
length(sg_all) #4230



in_vst <- SummarizedExperiment::assay(vst)[sg_all,FLT3.ITD_only$Sample]

dim(in_vst)
head(in_vst[,1:5]) #
```


```{r fig.height=4}
Cols <- c("FLT3.Groups_PoorRisk",  "FLT3.Groups_FavRisk","trisomy.8","Tissue","Batch")

cc <- colorCodes_aheatmap(FLT3.ITD_only[,Cols])
cc <- lapply(cc, function(x){ x[grep("All others", names(x))] <- "grey60"; return(x)})

# cc
par(mar=c(15,3,3,3))
col_check <- cc$FLT3.Groups_FavRisk
# barplot(rep(1,length(col_check)), col=col_check, names.arg = names(col_check), las=2)
```

Need to cite: devtools::install_github('cole-trapnell-lab/leidenbase') 


n_neighbors	
The size of local neighborhood (in terms of number of neighboring sample points) used for manifold approximation. Larger values result in more global views of the manifold, while smaller values result in more local data being preserved.

min_dist	
The effective minimum distance between embedded points. Smaller values will result in a more clustered/clumped embedding where nearby points on the manifold are drawn closer together, while larger values will result on a more even dispersal of points. 

```{r}
umap_bulk <- UMAP_workflow(TFIDF_Matrix = in_vst, 
                           scale_data=FALSE,
                           input_features = sg_all,
                            samples_vector = FLT3.ITD_only$Sample,
                            sample_info_df = FLT3.ITD_only,
                            Columns_for_Plots = Cols,
                            cc = cc, 
                            addl_color_vector=rainbow(n=40),
                            min_dist = 0.01,
                            n_neighbors=15,
                            k2=10,
                            res2=0.025)
```

```{r fig.height=10, fig.width=7}
umap_bulk$umap_2D_scatter
```


# Clustering with FLT3-ITD DEGs

```{r}
diagnostic_samples.aml <- diagnostic_samples %>% 
  filter(!is.na(FLT3.ITD.positive.)) %>% 
  set_rownames(.$Sample)

table(diagnostic_samples.aml$FLT3.ITD.positive.)
# table(diagnostic_samples.aml$Group)
# head(diagnostic_samples.aml$USI)


ITD.DE <- twoGroups_DEGs(expnData = cts_grch38,
                         clinData = diagnostic_samples.aml,
                         col = "FLT3.ITD.positive.", ref = "No",
                         percent.cutoff = 0.01,
                         anno = FALSE,
                         SkipPlots = TRUE)
```


```{r}
degs <- extract_DEGs(ITD.DE,anno = FALSE, geneLevel = FALSE)

dim(degs) # 1337    8
# head(degs)
# tail(degs)
# names(ITD.DE)
# saveRDS(ITD.DE,"TARGET_AML_FLT3-ITD+_vs_FLT3-ITD-_DE_results.RDS")
table()
```

## Heatmap

```{r}
 p <- diagnostic_samples.aml$FLT3.Groups_PoorRisk %>%
  set_names(diagnostic_samples.aml$Sample)


  dends_res_de <- dge_dendrograms(expnData = ITD.DE$DE$Voom$E,
                  pheno = p,
                  log=TRUE,
                  method = "ward.D2",
                  percent=0.01,
                  add.count=1,
                  createDGE = FALSE,
                  filterTopGenes = FALSE,
                  genelist = degs$gene)
  
  all(degs$gene %in% rownames(dends_res_de$TMMCPM))
  anno_deg <- DeGSEA::create_HA_Labs_Hmap(expn=dends_res_de$TMMCPM,
                                      geneList = degs$gene,
                                      cc = cc_heatmap2,
                                      CDE = diagnostic_samples.aml,
                                      cols = names(cc_heatmap2))


heatmap_degs <- ComplexHmap(mat = dends_res_de$TMMCPM, 
              name = "Z-Scores",
              scale=TRUE,
              dge_dendrograms.res = dends_res_de,
              hmap_anno_obj = anno_deg$annoColumn)
```

```{r}
# pdf("Figures/TARGET_AML_FLT3-ITD+_vs_FLT3-ITD-_from_DEGs.pdf", height = 10, width = 20)
heatmap_degs
# dev.off()
```

## UMAP

```{r}
#they are all the repetitive elements 
# degs$gene[!degs$gene %in% rownames(assay(vst))]

sapply(names(cc_heatmap), function(x) table(FLT3.ITD_only[[x]] %in% names(cc_heatmap[[x]])))
```

n_neighbors	
The size of local neighborhood (in terms of number of neighboring sample points) used for manifold approximation. Larger values result in more global views of the manifold, while smaller values result in more local data being preserved.

min_dist	
The effective minimum distance between embedded points. Smaller values will result in a more clustered/clumped embedding where nearby points on the manifold are drawn closer together, while larger values will result on a more even dispersal of points. 

```{r}
cc_degs <- cc_heatmap
# cc_degs[["AML_Subtype"]] <- c( "DEK-NUP214"= "#4DAF4A",
#                                "NUP98-NSD1" = "#377EB8",
#                                "AML, NOS" = "grey80",
#                                "No.Primary.Fusion" = "wheat2",
#                                "NUP98-KDM5A" = "royalblue1",
#                                "CBFB-MYH11" = "sienna",
#                                "RUNX1-RUNX1T1"="burlywood4",
#                                "KMT2A" = "khaki2")

cc_degs[["AML_Subtype"]] <- get_palette("Set1", length(unique(FLT3.ITD_only$AML_Subtype))) %>% 
  set_names(unique(FLT3.ITD_only$AML_Subtype))
cc_degs[["AML_Subtype"]]["DEK-NUP214"] <- "darkgreen"
cc_degs[["AML_Subtype"]]["KMT2A"] <- "khaki2"
cc_degs[["AML_Subtype"]]["NUP98-NSD1"] <- "skyblue2"

cc_degs[["AML_Subtype"]]["AML, NOS"] <-  "grey80"
cc_degs[["AML_Subtype"]]["No.Primary.Fusion"] <- "wheat1"
cc_degs[["AML_Subtype"]]["NUP98-KDM5A"] <- "mediumorchid3"
cc_degs[["AML_Subtype"]]["CBFB-MYH11"] <- "sienna"
cc_degs[["AML_Subtype"]]["RUNX1-RUNX1T1"] <- "burlywood3"


par(mar=c(10,5,5,5))
input <- cc_degs$AML_Subtype
barplot(rep(1,length(input)), col=input, names.arg = names(input), las=2)
```

```{r message=FALSE}
genes <- intersect(degs$gene,rownames(assay(vst))) #only the repetitve elements removed

FLT3_DEGs_UMAP  <- UMAP_workflow(TFIDF_Matrix = assay(vst)[genes,FLT3.ITD_only$Sample], 
                           scale_data=FALSE,
                           input_features = genes,
                            samples_vector = FLT3.ITD_only$Sample,
                            sample_info_df = FLT3.ITD_only,
                            Columns_for_Plots = names(cc_degs[c(2,1,3)]),
                            cc = cc_degs[c(2,1,3)], 
                            addl_color_vector=rainbow(n=40),
                            min_dist = 1e-3,
                            n_neighbors=3,
                            k2=10,
                            res2=0.015)

# nn=3
# min_dist=1e-3

```




```{r fig.height=12.5, fig.width=15}
# pdf("TARGET_AML_FLT3-ITD_vs_OtherAML_DEGs_UMAP_2D_scatter.pdf", height = 12.5, width = 15)
(FLT3_DEGs_UMAP$umap_2D_scatter$FLT3.Groups_PoorRisk + FLT3_DEGs_UMAP$umap_2D_scatter$AML_Subtype) / 
  (FLT3_DEGs_UMAP$cluster_plots1$scatter + FLT3_DEGs_UMAP$cluster_plots1$barplot)
# dev.off()
```

```{r}
FLT3_DEGs_3D <- scatter_plots_3d(umap_workflow_res = FLT3_DEGs_UMAP,
                                      Group_Column = "FLT3.Groups_PoorRisk", 
                                      Cols = c("Sample","FLT3.Groups_PoorRisk", "AML_Subtype","Mutations.Category", "Primary.Fusion","EFS.event.type.ID"),
                                      cc = cc_degs$FLT3.Groups_PoorRisk,
                                      blackbg = FALSE)
```


```{r}
FLT3_DEGs_3D
# htmlwidgets::saveWidget(FLT3_DEGs_3D, file="TARGET_AML_FLT3.ITD+_vs_OtherAML_byDEGs_minDist.1e-3_nn3_UMAP.html", selfcontained = TRUE) 
```

# Pathway Analysis 

Pathway Commons Search Pathways
Parameters:
q= [Required] a keyword, name, external identifier, or a Lucene query string.

page=N [Optional] (N>=0, default is 0). Search results are paginated to avoid overloading the search response. This sets the search result page number.

datasource= [Optional] filjsonter by data source (use names or URIs of pathway data sources or of any existing Provenance object). If multiple data source values are specified, a union of hits from specified sources is returned. For example, datasource=reactome&datasource=pid returns hits associated with Reactome or PID.

organism= [Optional] organism filter. The organism can be specified either by official name, e.g. "homo sapiens" or by NCBI taxonomy identifier, e.g. "9606". Similar to data sources, if multiple organisms are declared, a union of all hits from specified organisms is returned. For example 'organism=9606&organism=10016' returns results for both human and mouse. Note the officially supported species.

type= [Optional] BioPAX class filter (values). NOTE: queries using &type=biosource (or any BioPAX UtilityClass, such as Score, Evidence) filter won't not return any hits; use Entity (e.g., Pathway, Control, Protein) or EntityReference type (e.g., ProteinReference) instead.

Example Query: 
https://www.pathwaycommons.org/pc2/search?q=FGFR2&type=pathway
e.g. '/search.json'


```{r}
library(jsonlite)
```

```{r}
create_query_str <- function(search_term){
  url <- "https://www.pathwaycommons.org/pc2/"
  query_str <- paste0(url, paste0("search.json?q=", search_term, "&organism=9606&type=pathway"))
  
  res <- fromJSON(txt=query_str) 
  res[[3]] <- res[[3]] %>%
    mutate(search_term=search_term) %>% 
    select(1:2, search_term, everything())
  
  return(res)
}
```

FLT3-ITD Pathways 

The FLT3 cytoplasmic domain physically associates with the p85 subunit of phosphoinositol-3-kinase (PI3K), Ras GTPase [RAS], phospholipase C-γ [PLCG], Shc, growth factor receptor-bound protein (Grb2) and Src family tyrosine kinase, and results in the phosphorylation of these proteins [38]. These actions affect the activation of further downstream PI3K/protein kinase B (Akt) and mitogen-activated protein kinase (MAPK) pathways 

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3076284/

https://www.nature.com/articles/s41598-019-48687-z


```{r}
searches <- c("RAS","PLCG", "SHC", "GRB2", "SRC", "AKT", "MAPK","STAT5", "RAC1","LIG3","FOXO3", "CEBPA", "CXCR4", "PIM1")
```

```{r}
pathways <- purrr::map(searches, create_query_str)
pathways <- purrr::map_dfr(pathways, `[[`, 3) %>% 
  distinct()

head(pathways)
# dim(pathways) #521   9
# write.csv(select(pathways, uri:name),"TARGET_AML_FLT3-ITD_literature_textmining_associated_pathways.csv", row.names = FALSE)
```


Pathway Commons GET
Parameters:
uri= [Required] valid/existing BioPAX element's absolute URI (for utility classes that were "normalized", such as entity references and controlled vocabularies, it is usually an Identifiers.org URL. Multiple identifiers are allowed per query, for example, 'uri=http://identifiers.org/uniprot/Q06609&uri=http://identifiers.org/uniprot/Q549Z0' See also note about URIs and IDs.
format= [Optional] output format (values)
pattern= [Optional] array of built-in BioPAX patterns to apply (SIF types - inference rule names; see output format description) when format=SIF or TXT is used; by default, all the pre-defined patterns but neighbor-of apply.
subpw= [Optional] 'true' or 'false' (default) - whether to include or skip sub-pathways when we auto-complete and clone the requested BioPAX element(s) into a reasonable sub-model

Example: 
https://www.pathwaycommons.org/pc2/get?uri=http://identifiers.org/reactome/R-HSA-201451
https://www.pathwaycommons.org/pc2/get?uri=FGFR2&format=gsea

```{r}
get_pathways_gmt <- function(pathway_url){
  url <- "https://www.pathwaycommons.org/pc2/"
  request <- paste0("get?uri=",pathway_url,"&format=gsea&subpw=false")
  request <- paste0(url,request)

  paths_gmt <- RCurl::getURL(request)
  
  # edit so that the name is the first entry. so when read.gmt() is used the name, and not he URL is used as the pathway identifier.
  name <- gsub("^http.+\\\t(name:\\s.+)\\s\\[.+","\\1",paths_gmt)
  paths_gmt <- paste(name, paths_gmt, sep="\t")
  paths_gmt <- gsub("\\\tname:", "; name:", paths_gmt)
}
```

```{r}
tictoc::tic()
pathway_reqs <- purrr::map(pathways$uri, get_pathways_gmt)
tictoc::toc() #

length(pathway_reqs) #521
pathway_reqs <- unlist(pathway_reqs)
idx <- grep("^\\\t", pathway_reqs)
pathway_reqs <- pathway_reqs[-idx]
length(pathway_reqs) #454

#save the file to a gmt
# cat(unlist(pathway_reqs),
#     file = paste0("Pathway_Commons_v12_pathways_",Sys.Date(),".gmt"),
#     append = FALSE, sep="")
```

```{r}
FLT3_pathways_gmt <- read.gmt("Pathway_Commons_v12_pathways_2021-10-13.gmt")

length(FLT3_pathways_gmt)
# head(FLT3_pathways_gmt)
```



## GSVA 

```{r}
library(rhdf5)
# BiocManager::install("GSVA")
library(GSVA)
```

```{r}
# logCPM
gsva.res.all <- gsva(expr = logCPM,
                 gset.idx.list = FLT3_pathways_gmt,
                 # annotation=,
                 method="ssgsea",
                 kcdf="Gaussian",
                 parallel.sz=2, 
                 mx.diff=TRUE,
                 abs.ranking=FALSE, 
                 tau=1,
                 min.sz=5,
                 max.sz=400,
                 verbose=TRUE)


head(gsva.res.all[,1:5])
# dim(gsva.res.all) #423 1551
# write.csv(gsva.res.all, "GSEA/TARGET_AML_FLT3-ITD_ssGSEA_FLT3_associated_pathways.csv")
```

```{r}
pathways <- read.csv("TARGET_AML_FLT3-ITD_literature_textmining_associated_pathways.csv")
gsva.res.all <- read.csv("GSEA/TARGET_AML_FLT3-ITD_ssGSEA_FLT3_associated_pathways.csv")
# 
# 
dim(gsva.res.all) 
# head(gsva.res.all[,1:5])
```

```{r}
poor_vs_others <- FLT3.ITD_only %>% 
  filter(grepl("High risk|All others", FLT3.Groups_FavRisk)) %>% 
  mutate(FLT3.Groups_FavRisk=as.factor(FLT3.Groups_FavRisk)) %>% 
  set_rownames(.$Sample)

DE_poor_vs_others <- gsva_DE(gsva_matrix = gsva.res.all[,poor_vs_others$Sample],
                              clinData=poor_vs_others,
                              col="FLT3.Groups_FavRisk",
                              p.value = 1)
```

```{r}
# DE_poor_vs_others$contrast
# table(DE_poor_vs_others$fit$design[,"Ref"])
# table(DE_poor_vs_others$fit$design[,"Comparitor"])

DE_poor_vs_others_res <- read.csv("GSEA/TARGET_AML_HighRisk_FLT3-ITD_vs_allOthers_FLT3-ITD_ssGSEA_pathways.csv")

# DE_poor_vs_others_res <- DE_poor_vs_others$gene_sets %>% 
#   filter(adj.P.Val < 0.05)

dim(DE_poor_vs_others_res)
# View(DE_poor_vs_others_res)
# write.csv(DE_poor_vs_others_res, "GSEA/TARGET_AML_HighRisk_FLT3-ITD_vs_allOthers_FLT3-ITD_ssGSEA_pathways.csv", row.names = FALSE)
```

```{r}
top_paths <- DE_poor_vs_others_res %>% 
  filter(adj.P.Val < 0.01) %>% 
  mutate(Dir=ifelse(logFC > 0 , "up","zdown")) %>% 
  mutate(name=gsub("name: ", "", ID)) %>% 
  left_join(., select(pathways,uri:name),
            by="name") %>% 
  select(GeneSet:ID, search_term, everything())


top_paths_up <- head(top_paths)
top_paths_dn <- tail(top_paths)

# FLT3_pathways_gmt[top_paths_up$ID]
```

```{r}
top_paths_up
top_paths_dn
```



### Favorable

```{r}
fav_vs_others <- FLT3.ITD_only %>% 
  filter(grepl("Low risk|All others", FLT3.Groups_PoorRisk)) %>% 
  mutate(FLT3.Groups_PoorRisk=as.factor(FLT3.Groups_PoorRisk)) %>% 
  set_rownames(.$Sample)

table(fav_vs_others$FLT3.Groups_PoorRisk)
# levels(fav_vs_others$FLT3.Groups_PoorRisk)

DE_fav_vs_others <- gsva_DE(gsva_matrix = gsva.res.all[,fav_vs_others$Sample],
                  clinData=fav_vs_others,
                  col="FLT3.Groups_PoorRisk",
                  p.value = 1)
```

```{r}
DE_fav_vs_others_res <- read.csv("TARGET_AML_LowRisk_FLT3-ITD_vs_allOthers_FLT3-ITD_ssGSEA_pathways.csv") %>% 
  mutate(name=gsub("name: ", "", ID)) %>% 
  left_join(., select(pathways,uri:name),
            by="name") %>% 
  select(GeneSet:ID, search_term, everything())

# DE_fav_vs_others_res <- DE_fav_vs_others$gene_sets %>% 
#   filter(adj.P.Val < 0.05) s

# write.csv(DE_fav_vs_others_res, "TARGET_AML_LowRisk_FLT3-ITD_vs_allOthers_FLT3-ITD_ssGSEA_pathways.csv", row.names = FALSE)
```

```{r}
DE_fav_vs_others_res
```


# Session Information

```{r}
sessionInfo()
```

